---
title: "Introduction to R, part B"
author: "cs4821-s17"
---

[IntroA](introA.html) | [IntroB](introB.html) | [IntroC](introC.html) | [IntroD](introD.html)


## Working with Data 

R has provided a number of built-in data sets.  A list of data sets available can be viewed with:

```{r, eval=FALSE}
data()
```


## First Installation of Library

Let's explore using R with the _mpg_ data set available in the _ggplot2_ library.  If you have not yet installed this library on your computer/account then you will get the following error: `Error: there is no package called 'ggplot2'`

The first installation is down with the following command:

```{r eval=FALSE}
install.packages('ggplot2')
```

Once a package has been installed, then when you desire to use its functions, the library only needs to be loaded. 

```{r}
library(ggplot2)
data(mpg)
```

## Data Set Functions

Information about the data set may be viewed via in the Files/Plot/Packages/Help Panel of RStudio
```{r, eval=FALSE}
help(mpg)
```

A basic summary of the variables can be assessed by
```{r}
summary(mpg)
```

The numeric variables are listed with their minimum, maximum, 1st Quartile (25th percentile), 3rd Quartile (75th percentile), mean and median values.  Nominal variables show the frequency of each attribute values.

An alternative view of the data can be found by
```{r}
str(mpg)
```

The dimensions of the data (rows, columns) can be found by
```{r}
dim(mpg)
```

To look at the raw data, the top of the data matrix can be peeked at with
```{r}
head(mpg)
```

## Indexing (Subscripting)

Any part of the data matrix may be viewed via indexing (also referred to as subscripting)
```{r}
mpg[100:109,]
```

The data columns may be accessed by numeric indexing (the 8th - 11th variables) or by name ('cty', 'hwy', 'fl', 'class')
```{r}
mpg[80:85,8:11]
mpg[80:85,c("cty", "hwy", "fl", "class")]
```

A single data frame may be indexed
```{r}
mpg$hwy[1:20]
```

Data may be accessed by indicated items to exclude
```{r}
mpg[80:85,-1]
```

```{r}
mpg[80:85,-c(1,2,3,4,5)]
```


## Basic Plots

Let's try a little visualization:
```{r fig.width=7, fig.height=5}
plot(mpg[,c("displ", "cty")])
```

Or a box plot
```{r fig.width=7, fig.height=5}
boxplot(mpg$hwy ~ mpg$year)
```

