---
title: "Introduction to R, part A"
author: "cs4821-s17"
---

[IntroA](introA.html) | [IntroB](introB.html) | [IntroC](introC.html) | [IntroD](introD.html)



## Installation and Setup

[RStudio](http://www.rstudio.com/ide/) was used to construct the example activity.  The examples will be described and made available using [R Markdown](http://www.rstudio.com/ide/docs/r_markdown) (integrates [Markdown](http://daringfireball.net/projects/markdown), a simple formatting syntax for web pages, with R scripts) and [Sweave](http://www.statistik.lmu.de/~leisch/Sweave/) (integrates LaTeX, a document formatting syntax, with R scripts).  R Studio allows easy use and creation of these documents, but requires the [knitr](http://yihui.github.com/knitr/) R package.

The knitr package can be installed via:
```{r, eval=FALSE}
install.packages('knitr', dependencies=TRUE)
```

On the MTU lab machines, it will first try to install the package to the network drive, then ask to install the package to a local library (accept this choice).


# Basic R Usage

The following examples go through using several of R's basic functions

## Arithmetic Operations

```{r}
5 + 4
```

The output of this arithmetic operation is the number `9`.  The ## symbol is used to mark that this is an output.  In the R console the double hash does not appear.  The [1] states that 9 is on row 1 (the only row for this object, more on objects next).

```{r}
(10+13)/14.2
```

## Objects

Two main concepts exist in R: objects and functions.  Objects associate a name with a storage space; everything in R is stored as an object.  Content may be stored in objects using assignment `<-`.
 
Assignment may use an equal sign `=` but is not recommended.

```{r}
x <- 5
x
```

```{r}
y <- x^2
y
```

```{r}
sqrt(x)
```

Every object created stays in memory until deleted.  List all objects in memory using `ls()` or `objects()`.  Objects may be removed with `rm()`

```{r}
ls()
```

```{r}
rm(x)
```

### Object Types

Objects can have different types. There are four basic types: _characters_, _logical_, _numeric_, and _complex_.  The code above creates numeric objects for `x` and `y`.  The type of an object can be found with the `class()` function. 

```{r}
class(y)
```

Character objects can be created by enclosing a string inside single or double quotes.  

```{r}
a <- "Hello World"
a
```


## Vectors

The basic data object is a vector.  The objects `x` and `y` from above are vectors containing a single element.

Vectors store elements of the same type.  

A logical object can be created by assigning values of `T`, `F`, `TRUE`, or `FALSE`.

```{r}
b <- T
b
```

The length of a vector is assessed with the function `length()`

```{r}
length(y)
```

Vectors can be created with the concatenation function, `c()`, and can contain elements of the same type.  The vectors `v1` and `v2` hold _numeric_ objects and `v3` contains _character_ objects. 

```{r}
v1 <- c(1, 2, 3, 5)
v1
length(v1)
v2 <- c(2, 3, 5, 4)
v3 <- c("apple", "banana", "cherry", "date")
```

### Joining Vectors

Vectors can be combined using the `cbind` and `rbind` functions which binds vectors together as columns or as rows. 

```{r}
v4 <- cbind(v1, v2)
v4
```

```{r}
v5 <- rbind(v1,v2)
v5
```

```{r}
v6 <- cbind(v1,v3)
v6
```

Note, the matrix object created has only one type, _character_, the numbers are shown within quotes.


## Functions

Functions do things to objects, where functions may only run on objects of the correct type. 

```{r}
mean(v3)
```

Functions also work on vectors

```{r}
sum(v1)
```

```{r}
d <- sqrt(v1^2 + v2^2)
d
```

Vectors are indexed using square brackets `[ ]` starting with 1
```{r}
v1[1]
```

## Data Frames

When a matrix of data is desired that contains objects of different types a data frame is used.  

```{r}
df <- data.frame(v1, v3)
df
```

Data frames have column names associated with each column, in this case the name of the vector object.

```{r}
names(df)
```

The rows of a data frame may also have names (when not specified the row names default to numeric values). 

```{r}
row.names(df)
```

You can assign other values to the row names. 

```{r}
row.names(df)  <- c("One", "Two", "Three", "Four")
row.names(df)
```

## Factors

Factors are a way to store categorical (nominal) data.  A factor has _levels_, possible values, to take. 

```{r}
g <- c("f", "m", "m", "f", "f", "f", "m")
g
```

```{r}
g <- factor(g)
g
```

Look at distribution of possible values

```{r}
table(g)
```

A cross-tabulation of several factors can also use the `table()` function

```{r}
m <- factor(c("cs", "math", "cs", "cs", "math", "math", "cs"))
table(m,g)
```

## Sequences

Construct a sequence of values

```{r}
x <- 1:10
x
```

Decreasing sequences may also be created. 

```{r}
y <- 5:0
y
```

Other sequences may be constructed with the `seq()` function

```{r}
seq(0, 20, by=2)
```

```{r}
seq(1, -3, -0.5)
```

Sequences may also be generated using repeating patterns

```{r}
rep(3,4)
```

```{r}
c(rep(2,3), rep(4,2))
```

