---
title: "Introduction to R, part D"
author: "cs4821-s17"
---

[IntroA](introA.html) | [IntroB](introB.html) | [IntroC](introC.html) | [IntroD](introD.html)



The following examples go through using several of R's basic functions.

Load the _mpg_ data set.

```{r}
library(ggplot2)
data(mpg)
```

## Sorting

Sort values using the `sort` function
```{r}
x <- mpg$hwy[1:10]
x
sort(x)
```

The ordering of a vector may also be found
```{r}
y <- 1:10
o <- order(x)
x[o]
```

```{r}
y[o]
```

## Functions

A new function can be defined by assigning the function to the object name.  Let's create a function to calculate standard error.

$$ \text{standard error} = \sqrt{\frac{s^2}{n}} $$

```{r se}
se <- function(x) {
  v <- var(x)
  n <- length(x)
  return(sqrt(v/n))
}
se(c(45, 2, 5, 8, 65, 4))
```

A function may be used to calculate basis statistics on a vector:

```{r basicstat}
basic.stats <- function(x, more=F) {
  stats <- list()
  
  clean.x <- x[!is.na(x)]
  
  stats$n <- length(x)
  stats$nNAs <- stats$n - length(clean.x)
  
  stats$mean <- mean(clean.x)
  stats$std <- sd(clean.x)
  stats$med <- median(clean.x)
  if (more) {
    stats$skew <- sum(((clean.x-stats$mean)/stats$std)^3) / 
      length(clean.x)
  }
  unlist(stats)
}
basic.stats(c(40, 2, 3, 5, 10, 55, 8, -400))
```

```{r}
basic.stats(c(40, 2, 3, 5, 10, 55, 8, -400), more=T)
```

This example also introduced some program control namely `if` functions

## Program Control

A simple example using _for_ loops

```{r}
for (i in 1:10) {
  print(i^2)
}
```

```{r}
for (i in c("red", "green", "blue")) {
  print(i)
}
```

Note, R like vector has many functions that work on vectors. Therefore, vectorized version of calculations and operations are often preferred over loops to ensure efficient code.

```{r}
i <- 0
while (i < 10) {
  print(i)
  i <- i + 1
}
```

## Plotting and Saving Results

Let's look at a basic scatter plot

```{r}
x = c(1:40)
y = x*2 + 10 + rnorm(40)*3
plot(x, y, pch=21, bg="red",
     main="A Sample Plot", xlab="dep. variable", ylab="ind. variable")
```

Plots can be saved to several image formats: _png_, _jpeg_, _bmp_, _tiff_.  The plot will be saved in a file: "R-introD-fig1.png"

```{r, include=FALSE}
png("R-introD-fig1.png")
plot(x, y, pch=21, bg="red",
     main="A Sample Plot", xlab="dep. variable", ylab="ind. variable")
dev.off()
```


