---
title: "Introduction to R, part C"
author: "cs4821-s17"
---

[IntroA](introA.html) | [IntroB](introB.html) | [IntroC](introC.html) | [IntroD](introD.html)


## Matrices and Arrays

You may vectorize a series of numbers into a matrix, via setting the dimensions.

```{r}
m <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
m
```

```{r}
dim(m) <- c(2, 5)
m
```

A matrix may also be created using the `matrix()` function

```{r}
m2 <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), 2, 5)
m2
```

Matrices may be joined by columns `cbind()` or by rows `rbind()`

```{r}
cbind(m[,c(2,3)], m2[,c(5,3)])
```

```{r}
rbind(m[2,c(3,4,5)], c(1,2,3))
```


## Statistical Measures

Load the _mpg_ data set

```{r}
library(ggplot2)
data(mpg)
```

Many statistical measure can be assessed for numeric data: mean, median, min, max, and range of a variable 

```{r}
mean(mpg$hwy)
```

```{r}
median(mpg$hwy)
```

```{r}
min(mpg$hwy)
max(mpg$hwy)
```

```{r}
range(mpg$hwy)
```

The standard deviation, variance, and correlation may also be calculated

```{r}
sd(mpg$hwy)
var(mpg$hwy)
cor(mpg$hwy,mpg$cty)
```

## Random Numbers

Create vectors of random numbers from standard distributions.  
* Uniform distribution
* Gaussian distribution
* Exponential distribution
* Poisson distribution

```{r}
runif(10)
```

Scale the values to be between 0 and 100

```{r}
runif(10, 0, 100)
```

Sample a standard Gaussian distribution, mean = 0, std = 1

```{r}
x <- rnorm(100)
mean(x)
sd(x)
```

Sample a non-standard Gaussian distribution with mean=5, and std = 3
```{r}
y <- rnorm(100, 5, 3)
mean(y)
sd(y)
```

```{r}
rexp(5)
```

```{r}
rexp(10,1)
rexp(10,10)
```

## Common R functions

* Absolute value: `abs`
```{r}
abs(-5)
```
* Sign function: `sign`
```{r}
sign(-2)
```
* Square root: `sqrt`
```{r}
sqrt(2)
```
* Logarithms: `log`, `log2`, `log10`, `logb`
```{r}
log(100)
```
```{r}
log2(16)
```

* Exponential: `exp`
```{r}
exp(1)
```
* Floor: `floor`
```{r}
floor(8.543)
```
* Ceil: `ceil`
```{r}
ceiling(8.543)
```
* Truncate: `trunc`
```{r}
trunc(8.543)
```
* Round: `round`
```{r}
round(8.543)
```
* Round to significant digits: `signif`
```{r}
signif(3.14159,3)
```
* Trigonometric functions: `sin`, `asin`, `cos`, `acos`, `tan`, `atan`
```{r}
c(sin(45), cos(45), tan(45))
```
* Mod operations: `%/%`,  `%%`
```{r}
c(10%/%3,  10%%3)
```

