\documentclass[12pt, letterpaper]{report}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,bottom=1in,top=1in]{geometry}
\author{Taylor B. Morris}
\title{Case Study Report}

\begin{document}
\makeatletter
{\huge\noindent\@title\large\\\@author\\\@date}
\makeatother

A pair of computer scientists from University of Kansas plagiarised several
works, taking for their own statements up to and including an entire opening
statement. Visvanathan and Lushington were discovered to have plagiarised
statements and research in three of their papers. Sadly, their plagiarism
was not detected by automated detection software at time of submission, but
much later on by the scientists whom they had plagiarised. These scientists
had noticed that another work of theirs had been officially cited, and they
went out to read the paper that cited their work. When they discovered the
plagiarism, they complained to the journal.

While Visvanathan was technically the first author and the submitter of the
plagiarised pieces, Lushington - Visvanathan's supervisor - was held
similarly responsible. This is because a student had brought the plagiarism
to his attention before it was officially reported, and Lushington chose not
to report the misconduct. 

While it is unclear why either committed the misconduct, Visvanathan claimed
to be taking the fall for an unnamed international student. Lushington chose
not to comment on the misconduct, but it is likely he did not report it due
to his standing of being a listed author on each of the three papers.

In the instance of the third paper, Visvanathan and Lushington were on their
way to a conference in Sweden when their misconduct was discovered. The
poster they were about to present on said paper was copied word-for-word
from someone else's paper. 

In this case, the research which was plagiarised was already published in
research journals. These researcher who had their work plagiarised lost
citations. IEEE formally retracted the papers from Visvanathan and
Lushington, and researchers who referred to the pieces written by 
Visvanathan and Lushington in their pubished or unpublished papers had to remove references to the retracted papers.


\end{document} 
